
```{r setup_cleaning_preparation, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```


```{r}
## Cleaning data

library(dplyr)

# Not informative predictors
DF <- DF %>%
  select(-AnimalID, -DateTime)

str(DF)

```

# Feature Engineering

## Names

### NAs

```{r}

# How many names are missing. SO much data records will disapear during training a model if we don't do anything with NA.
sum(is.na(DF$Name))/nrow(DF)

DF <- DF %>%
  mutate(Name = ifelse(is.na(Name), "not_known", Name))

```


### How many unique names. 

Each level of a factor will create a new variable. Do we need so many new variables?

Let us replace Name by a status of popular or not

Take popular names from

* http://www.youpet.com/cat-names/
* http://www.youpet.com/dog-names/



```{r}

length(unique(DF$Name))

pop_names_cat <- read.csv("data/top_100_cat_names.txt") 
pop_names_dog <- read.csv("data/top_100_dog_names.txt")   

library(stringr)
pop_names_cat <- str_trim( as.character(pop_names_cat[, 2]) )
pop_names_dog <- str_trim( as.character(pop_names_dog[, 2]) )

DF <- DF %>%
  mutate(PopNameD = AnimalType == "Dog" & Name %in% pop_names_dog) %>%
  mutate(PopNameC = AnimalType == "Cat" & Name %in% pop_names_cat) %>%
  mutate(PopName = ifelse(PopNameC | PopNameD, "yes", "no")) %>%
  mutate(PopName = ifelse(Name == "not_known", "not_known", PopName)) %>%
  select(-PopNameC, -PopNameD, -Name)
  
table(DF$PopName)

```

## Breed

```{r}

#   Breed
length(unique(DF$Breed))
head(DF$Breed)

DF <- DF %>%
  mutate(BreedSH = factor( str_detect(Breed, "Shorthair")) ) %>%
  mutate(BreedL = factor( str_detect(Breed, "Long"))) %>%
  mutate(BreedD = factor( str_detect(Breed, "Domestic"))) %>%
  mutate(BreedMix = factor( str_detect(Breed, "Mix"))) %>%
  mutate(Breed2 = factor( str_detect(Breed, "/"))) %>%
  select(-Breed)

table(DF$BreedSH)
table(DF$BreedL)
table(DF$BreedD)
table(DF$BreedMix)
table(DF$Breed2)

head(DF)

```


##Color

```{r}
length(unique(DF$Color))

#find unique words for color
un_col <- str_replace_all(DF$Color, "[^[:alpha:]]", " ")  %>%
  word() %>%
  table() 

un_col
un_col <- un_col[!un_col < 50]   
un_col_names <- names(un_col)

for(nnn in un_col_names){
  DF[, nnn] <- as.factor(str_detect(DF$Color, nnn))
}

DF$Color <- NULL

```

## AgeuponOutcome

```{r}

DF <- DF %>%
    mutate(AgeOnOutcome = 
      as.numeric( str_extract(AgeuponOutcome, "[:digit:]+") ) *
         str_detect(AgeuponOutcome, "year")*360 +
      as.numeric( str_extract(AgeuponOutcome, "[:digit:]+") ) *
        str_detect(AgeuponOutcome, "month")*30 +
      as.numeric( str_extract(AgeuponOutcome, "[:digit:]+") ) *
        str_detect(AgeuponOutcome, "week")*7 + 
      as.numeric( str_extract(AgeuponOutcome, "[:digit:]+") ) *
       str_detect(AgeuponOutcome, "day")*1)

DF$AgeuponOutcome <- NULL

#View(DF)

```

## non-linear features

Let us add non-linear features for time.

```{r}
DF <- DF %>%
    mutate(AgeOnOutcome2 = AgeOnOutcome^2, 
           AgeOnOutcome12 = AgeOnOutcome^(1/2), 
           AgeOnOutcome3 = AgeOnOutcome^3, 
           AgeOnOutcome13 = AgeOnOutcome^(1/3))

```



## Other predictors

THis can't be used in prediction and is not helpfull right now. Remove this for now..

```{r}
DF$OutcomeSubtype <- NULL

```

```{r}
length(unique(DF$SexuponOutcome))
```

## Convert factors to factors

```{r}
DF[, 1:4] <- lapply(DF[, 1:4], as.factor )

```


# Dimension

```{r}
dim(DF)
str(DF)
sapply(DF, is.factor)

```


# Create Binary outcome

Create data frame for binary classification

```{r}
# alive or not
DF_B <- DF %>%
  mutate(OutcomeType = OutcomeType %in% c("Adoption")) %>%
  mutate(OutcomeType = ifelse(OutcomeType, "yes", "no")) 

# Try later
# alive or not. Classes will be unbalanced. Important to observe high accuracy even with a non-predicting model... just becaus of no-information rate
#DF_B <- DF %>%
#  mutate(OutcomeType = OutcomeType %in% 
#           c("Adoption", "Return_to_owner", "Transfer")) %>%
#  mutate(OutcomeType = ifelse(OutcomeType, "yes", "no")) 



# outcome var should not be logical TRUE or FALSE, and should not be number
#otherwise caret gives error for twoClassSumamry
# like At least one of the class levels are not valid R variables names
# because TRUE and FALSE can't be names of col.


#View(DF_B)
table(DF_B$OutcomeType)

```

