---
title: "Drink Recommender"
author: "Matt Moocarme"
date: "June 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      cache = T,
                      message = FALSE,
                      comment = '',
                      echo = F)
```

```{r}
library(jsonlite)
library(readr)
library(dplyr)
library(tidyr)
library(stringr)


margarita<- as.data.frame(fromJSON('http://www.thecocktaildb.com/api/json/v1/1/search.php?s=margarita'))
head(margarita)

by_id <- as.data.frame(fromJSON('http://www.thecocktaildb.com/api/json/v1/1/lookup.php?i=15112'))
str(by_id)

#database=data.frame()
#for(i in 10000:20000) {
# link="http://www.thecocktaildb.com/api/json/v1/1/lookup.php?i="
#  z=paste(link,i,sep="")
#  data=as.data.frame(fromJSON(z))
#  database=rbind(database,data)
#  print(i)
#}

database <- read_csv('data/database.csv')
str(database)
dim(database)


db1 <- database %>% select(cocktail.name = drinks.strDrink, 
                           drinks.strIngredient1:drinks.strIngredient7,
                           drinks.strMeasure1:drinks.strMeasure7)

```

```{r}
nas <- c('',' ','/t','\n')
db1[db1 == ''] = NA
db1[db1 == ' '] = NA
db1[db1 == '/t'] = NA
db1[db1 == '\n'] = NA

sum(is.na(db1))

db1[is.na(db1)] = 0
sum(is.na(db1))

db <- as.data.frame(lapply(db1, as.character), stringsAsFactors = F)

db_ingridient <- db %>% select(cocktail.name:drinks.strIngredient7) %>% 
  gather(Ingredient.number, Ingredient, drinks.strIngredient1:drinks.strIngredient7, -cocktail.name) %>%
  select(-Ingredient.number)
db_measurement <- db %>% select(cocktail.name, drinks.strMeasure1:drinks.strMeasure7) %>%
  gather(Measurement.number, Measurement, drinks.strMeasure1:drinks.strMeasure7, -cocktail.name) %>% 
  select(-Measurement.number)

db_join <- db_measurement %>% bind_cols(select(db_ingridient, Ingredient)) 

db2 <- db %>% gather(Ingredient.Number,Ingredient,drinks.strIngredient1:drinks.strIngredient7, -cocktail.name) %>%
  gather(Measurement.Number, Measurement, drinks.strMeasure1:drinks.strMeasure7, -cocktail.name)

db2 <- db2 %>% filter(substr(Measurement.Number, nchar(Measurement.Number), nchar(Measurement.Number))== substr(Ingredient.Number, nchar(Ingredient.Number), nchar(Ingredient.Number)))

db3 <- db2 %>% select(-Ingredient.Number, -Measurement.number) 

db4 <- db_join %>% filter(Ingredient != '0') %>%
  group_by(Ingredient) %>% 
  summarise(Total.count = n()) %>% 
  arrange(desc(Total.count))

head(db4, 20)


db5 <- db_join %>% filter(Ingredient != '0') %>%
  group_by(Measurement) %>% 
  summarise(Total.count = n()) %>% 
  arrange(desc(Total.count))

head(db5, 20)

meas <- unique(db3$Measurement)
(filter(meas, Measurement %in% c('oz','ml')))

```
```{r}
# for mac users. remove white spaces.
db_tidy <-mutate(db3,
               amount = str_trim(db3$Measurement, side = "both"))# separate numbers into its own column. # We will create a new column in each step to see the reduction in anomalities. # create a new column with numbers with punctuation.
db_tidyz <- db_tidy %>% mutate(number=gsub("[^[:digit:],^[:punct:] ]", "", Measurement)) #%>% #head()
 
# Trim both sides from whitespaces
db_tidyz <- db_tidyz %>% mutate(numbe=str_trim(number,side="both"))
 
# replace double and triple and so on whitspaces with single whitespace
 
db_tidyz <- db_tidyz %>% mutate(numb=gsub("\\s+"," ",numbe)) #%>% #head()
 
# Select all numbers that matching the following four patterns
db_tidyz <- db_tidyz %>% mutate(num=str_extract(number,"[:digit:]+\\s+[:digit:]+[:punct:]+[:digit:]|[:digit:]+[:punct:]+[:digit:]+|[:digit:]+|[:digit:]+[:punct:]+[:digit:]+[:digit:]+")) #head()# Replace all whitespaces with + sign
db_tidyz <- db_tidyz %>% mutate(nu=str_replace_all(num," ","+"))#head()# Separate everything that has "-" into two columns
db_tidyz <- db_tidyz %>% separate(nu,c("nu","nuu"), sep="-") # head()# Select text only (with white spaces)  
db_tidyz <- db_tidyz %>% mutate(unit=gsub("[[:digit:]|[:punct:]]","",amount)) # head()
 
# Trim both sides from whitespaces
db_tidyz <- db_tidyz %>% mutate(unit=str_trim(unit,side="both")) 
 
#replace double and triple and so on whitspaces with single whitespace
db_tidyz <- db_tidyz %>% mutate(unit=gsub("\\s+"," ",unit))# set NAs to 1
db_tidyz[is.na(db_tidyz)]<-1#Check unique values for number, numbe, num and unit
unique(db_tidyz$number)
unique(db_tidyz$numbe)
unique(db_tidyz$numb)
unique(db_tidyz$num)
unique(db_tidyz$nu)
unique(db_tidyz$unit)

```

```{r ingrindient_to_ml}
# Select only cocktail name, ingredient, nu number and unit column
db_tidyx <- db_tidyz %>%
   select(cocktail.name,Ingredient,nu, unit)

db_tidyx %>% head()

# Replace units with proper conversion to mL
db_tidyx <- db_tidyx %>%
   mutate(unit=str_replace(unit,"ozjamaican","oz")) %>%
   mutate(unit=str_replace(unit,"oz","29.5")) %>%
   mutate(unit=str_replace(unit,"shot","29.5")) %>%
   mutate(unit=str_replace(unit,"jigger","44.5")) %>%
   mutate(unit=str_replace(unit,"cup","257")) %>%
   mutate(unit=str_replace(unit,"tblsp","11.1")) %>%
   mutate(unit=str_replace(unit,"tsp","3.7")) %>%
   mutate(unit=str_replace(unit,"ts p","3.7")) %>%
   mutate(unit=str_replace(unit,"teaspoon","3.7")) %>%
   mutate(unit=str_replace(unit,"cl","10")) %>%
   mutate(unit=str_replace(unit,"dl","100")) %>%
   mutate(unit=str_replace(unit,"litre","1000")) %>%
   mutate(unit=str_replace(unit,"liter","1000")) %>%
   mutate(unit=str_replace(unit,"dash","0.9")) %>%
   mutate(unit=str_replace(unit,"splash","3.7")) %>%
   mutate(unit=str_replace(unit,"twist","15")) %>%
   mutate(unit=str_replace(unit,"twistof","15")) %>%
   mutate(unit=str_replace(unit,"can","355")) %>%
   mutate(unit=str_replace(unit,"cube","12")) %>%
   mutate(unit=str_replace(unit,"part","29.5")) %>%
   mutate(unit=str_replace(unit,"pint","473")) %>%
   mutate(unit=str_replace(unit,"glass","473"))

# Check if missing something like glass or a pint add it to the code on top
unique(db_tidyx$unit)

```
```{r dealing_with_fractions}


db_tidyc<-db_tidyx %>%
   # Select all numbers that matching the following patterns
   mutate(unit=str_extract(unit,"[:digit:]+[:digit:]+[:punct:]+[:digit:]|[:digit:]+[:punct:]+[:digit:]+|[:digit:]+|[:digit:]+[:punct:]+[:digit:]+[:digit:]+"))   #head()
   
   # Convert unit column to numeric
db_tidyc <- db_tidyc %>% mutate(unit=as.numeric(unit))  
   
   # replace comma , with dot .
db_tidyc <- db_tidyc %>% mutate(nu=gsub(",",".",nu))  
   # add left and right brackets for nu colum for proper parsing and evaluation
db_tidyc <- db_tidyc %>%    mutate(nu=sub("^", "(",nu))    #head()
db_tidyc <- db_tidyc %>%    mutate(nu=sub("$", ")",nu))  #%>%  head()

# set NAs to 1
db_tidyc[is.na(db_tidyc)]<-1
      
# parse and evaluate nu
m2 <- sapply(db_tidyc$nu ,function(x) eval(parse(text=x)))

str(m2)

# add m2 as a new column 
db_tidyc$nup<-m2

# compare nu and nup for sanity check 
head(db_tidyc,10)

# looks good

# multiply nup and unit for the measure
db_tidyc<- db_tidyc %>%
   mutate(measure=nup*unit)

#db_tidyc %>% head(10)

#  Select only cocktail name, ingridient and measure
db_clean_tidy <- db_tidyc %>%
   select(cocktail.name,Ingredient,measure)

db_clean_tidy %>% head(10)
```


```{r}
db_clean_tidy$row <- 1:nrow(db_clean_tidy)
db_spread <- db_clean_tidy %>% spread(Ingredient, measure) %>% select(-row)

db_spread[is.na(db_spread)]<-0

# Group by cocktail name
db_spread_comb <- db_spread %>%
   group_by(cocktail.name) %>%
   summarise_each(funs(sum))



# Check ingridients for a random cocktail
i=37
colnames(db_spread_comb)[which(db_spread_comb[i,]!=0)]

"salt" %in% colnames(db_spread_comb)

# Spread is ready! Let's save it!
write.csv(db_spread_comb,"db_spreaded.csv")
```

```{r}
dtf <- db_spread_comb[,-1]
km.out <- kmeans(dtf, 10, nstart = 20, iter.max = 50)
dclust <- data.frame(km.out$cluster, db_spread_comb$cocktail.name)

dtf <- as.data.frame(sapply(dtf, as.numeric))

xcorr<-cor(dtf)
xcorr[is.na(xcorr)]<-0

cor_mat<-xcorr
most_cor <- which(abs(cor_mat) > 0.75 & cor_mat != 1,
                 arr.ind=TRUE)
rownames(cor_mat)[7]
```

```{r}
angle <- function(x,y){
 dot.prod <- x%*%y
 norm.x <- norm(x,type="2")
 norm.y <- norm(y,type="2")
 theta <- acos(dot.prod / (norm.x * norm.y))
 as.numeric(theta/3.14*180)
}

# Prepare ingredient term matrix
dtf <- sapply(dtf, as.numeric)# Create dummy inner product matrix
sumi<-matrix(nrow=nrow(dtf),ncol=nrow(dtf))# Test if angle function works, answer is in degrees
angle(dtf[100,],dtf[220,])# Apply angle function on ingredient term matrix (without cocktail names), 

#row by row

#for (i in 1:nrow(dtf)){
#   for (j in 1:nrow(dtf)){
#   sumi[i,j]<-angle(dtf[i,],dtf[j,])
#   print(i)
#   print(j)
#   }
# }


# Replace NAs with 0
sumi[is.na(sumi)]<-0# Get histogram
hist(sumi)# Convert to dataframe set rows and columns names to the names of cocktails
sumidf<-as.data.frame(sumi)
colnames(sumidf)<-db_spread_comb$cocktail.name
rownames(sumidf)<-colnames(sumidf)# Replace NAs with 0s
sumidf[is.na(sumidf)]<-0# Save as db_innerproduct_matrix.csv
#write.csv(sumi,"cocktailz/db_innerproduct_matrix.csv")
## Explore inner products matrix```{r Explore_Inner_Products, eval=FALSE}
# Load Inner Products Matrix, set rownames the same as column names
#dfsumi<-read.csv("db_innerproduct_matrix.csv", header = TRUE)
dfsumi<-dfsumi[,-1] # take out row number colnames(dfsumi) <- db_spread_comb$cocktail.name[1:3156]
rownames(dfsumi)<-colnames(dfsumi)# set NAs to zero
dfsumi[is.na(dfsumi)] <- 0
# Choose Margarita
x<-dfsumi %>%
  select(Margarita)
rownames(x)<-colnames(dfsumi)
x<-sapply(x, as.numeric)y<-colnames(dfsumi)
hist(x[x<70])# Perform unsupervised kmeans clustering on inner products matrix
km.out.ip <- kmeans(dfsumi, 10, nstart =20,iter.max=50)# previous cluster
ip.x <-as.data.frame(km.out.ip$cluster)ip.x <- bind_cols(ip.x, as.data.frame(rownames(ip.x)))
```


